### 2020-12-08 paper/localpsf.tex
writing existing work

\subsection{Connection to existing methods}

In general, product-convolution approximations are approximations of the form \eqref{eq:product_convolution}, but where $\varphi_i$ and $w_i$ may be arbitrary functions. 
For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].

@@

@@

Geometry-aware weighting functions vs radial basis functions

@@

Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations, 
\begin{equation*}
y=f(m,u(m)) + \text{noise},
\end{equation*}
which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, 
\begin{equation}
\label{eq:state_pde}
	0=g(m,u).
\end{equation}
Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
%
%That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy measurements,
%\begin{equation*}
%	y = B(m,u) + \xi,
%\end{equation*}
%of a state variable, $u$, which depends on $m$ indirectly through the solution of a PDE,
%\begin{equation}
%\label{eq:state_equation}
%	0=F(m,u).	
%\end{equation}
%Here $\xi$ is the noise. The observations $y$ are known, while $m$, $u$, and $\xi$ are unknown. In the deterministic setting, the inverse problem of finding $m$ is commonly framed as a least squares minimization problem
%\begin{equation}
%\label{eq:minimization_problem}
%	\min_m \quad J_d(m) + R(m),
%\end{equation}
%where
%\begin{equation*}
%	J_d(m) := \frac{1}{2}\|y - B(m,u(m))\|_{\Gamma}^2
%\end{equation*}
%is a data misfit term, $\Gamma$ is the noise covariance, $R(m)$ is a regularization term, and $u(m)$ denotes the solution $u$ of the PDE \eqref{eq:state_equation} as a function of $m$, . 
%
%Approximation of the Hessian,
%\begin{equation*}
%	\mathcal{H} = \mathcal{H}_d + \mathcal{H}_R := \frac{d^2 J_d}{dm^2} + \frac{d^2 R}{dm^2},
%\end{equation*}
In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem 
\begin{equation}
\label{eq:minimization_problem}
	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
\end{equation}
where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term. Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods. Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$. 
We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs.

>>>>

\subsection{Existing work}

In general, product-convolution approximations take the form \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions. Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].

We build upon a line of work in which the convolution kernels, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources [CITE BIG LIST]. The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$. In [CITE NAGY], the impulse responses are sampled on a regular rectilinear grid, and interpolated with piecewise linear weighting functions, $w_i$, on a triangle mesh subordinate to that grid. Adaptive refinement of the grid was proposed in [CITE], and we demonstrated a practical algorithm for adaptive refinement in our previous work [CITE MY PAPER]. The issue with all of these existing methods, however, is computing each $\varphi_i$ requires one matrix vector product with $\mathcal{H}$, so that $r$ matrix-vector products are required. This is expensive when $\mathcal{H}$ is the Hessian in a PDE-constrained inverse problem, because each matrix vector product with $\mathcal{H}$ requires solving two PDEs. A key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product with $\mathcal{H}$, thereby greatly reducing the number of matrix-vector products with $\mathcal{H}$ that are required.

We do this by taking advantage of the local support property of $\mathcal{A}$.

SVIR low rank: sample columns of matrix

Pros and cons of different point picking strategies depend on the application of interest.

In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.

@@

H-matrices for Hessians: Ilonas paper with KAUST people

Demanet matrix probing can be interpreted as convolution-product.

Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.

@@

Geometry-aware weighting functions vs radial basis functions. Kriging.

@@

Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations, 
\begin{equation*}
y=f(m,u(m)) + \text{noise},
\end{equation*}
which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, 
\begin{equation}
\label{eq:state_pde}
	0=g(m,u).
\end{equation}
Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
%
%That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy measurements,
%\begin{equation*}
%	y = B(m,u) + \xi,
%\end{equation*}
%of a state variable, $u$, which depends on $m$ indirectly through the solution of a PDE,
%\begin{equation}
%\label{eq:state_equation}
%	0=F(m,u).	
%\end{equation}
%Here $\xi$ is the noise. The observations $y$ are known, while $m$, $u$, and $\xi$ are unknown. In the deterministic setting, the inverse problem of finding $m$ is commonly framed as a least squares minimization problem
%\begin{equation}
%\label{eq:minimization_problem}
%	\min_m \quad J_d(m) + R(m),
%\end{equation}
%where
%\begin{equation*}
%	J_d(m) := \frac{1}{2}\|y - B(m,u(m))\|_{\Gamma}^2
%\end{equation*}
%is a data misfit term, $\Gamma$ is the noise covariance, $R(m)$ is a regularization term, and $u(m)$ denotes the solution $u$ of the PDE \eqref{eq:state_equation} as a function of $m$, . 
%
%Approximation of the Hessian,
%\begin{equation*}
%	\mathcal{H} = \mathcal{H}_d + \mathcal{H}_R := \frac{d^2 J_d}{dm^2} + \frac{d^2 R}{dm^2},
%\end{equation*}
In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem 
\begin{equation}
\label{eq:minimization_problem}
	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
\end{equation}
where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term. Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods. Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$. 
We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
